---
title: "Other health indicators"
output: html_notebook
---

```{r}
library(dplyr)
library(RSQLite)
library(ggplot2)
library(broom.mixed)
library(tidyr)
library(brms)
library(bayesplot)
library(rstan)
library(multcomp)
library(coda)
library(tidybayes)

# options for rstan
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

In this notebook we'll examine the correlation between the neuroimaging derived phenotype scores and clinical scores in the UK biobank population.

```{r}
# create sqlite connection
ukbb.db <- DBI::dbConnect(SQLite(),
                          "../data/ukb40183.db")
```

```{r}
# define clinical score
threshold = 0.5

getdata <- function(clinscore, threshold){
  # pairs matching
  alldata <- dbGetQuery(ukbb.db,
                        sql(paste("SELECT clinicaldata.eid, clinicaldata.noncancer_code_selfreported,",
                                  "demogs.age, demogs.visit,",
                                  paste("clinicaldata", clinscore, sep = "."), " AS clinscore,",
                                  "ukbb_t1_phenotype.mean",
                                  "FROM clinicaldata",
                                  "JOIN ukbb_t1_phenotype, demogs ON clinicaldata.eid = ukbb_t1_phenotype.eid",
                                  "AND clinicaldata.eid = demogs.eid",
                                  "AND clinicaldata.visit = demogs.visit",
                                  "AND ukbb_t1_phenotype.eid = demogs.eid",
                                  "WHERE clinicaldata.t1_structural IS NOT ''",
                                  "AND demogs.age IS NOT ''",
                                  "AND", paste("clinicaldata", clinscore, sep = "."), "IS NOT ''")))

  alldata$adlike <- sapply(alldata$mean,
                           function(x, thr) if (x >=  thr) { return("Positive") } else {return("Negative")},
                           thr = threshold)
  
  alldata <- alldata %>%
    mutate(adlike = ifelse(noncancer_code_selfreported == "1263", "AD", adlike))
  
  alldata$adlike <- as.factor(alldata$adlike)
  alldata$clinscore <- as.numeric(alldata$clinscore)
  alldata$age <- as.numeric(alldata$age)
  
  # normalise age
  mean_age <- mean(alldata$age)
  sd_age <- sd(alldata$age)
  alldata$age <- (alldata$age - mean_age) / sd_age
  
  return(alldata)
}

```

# Other health indicators

## hand grip

### left
```{r}
clinscore = "hand_grip_left"
clinscore_name = "Left hand grip"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

### right
```{r}
clinscore = "hand_grip_right"
clinscore_name = "Right hand grip"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## waist circumference
```{r}
clinscore = "waist_circumference"
clinscore_name = "Waist circumference"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## hip circumference
```{r}
clinscore = "hip_circumference"
clinscore_name = "Hip circumference"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## BP
### systolic - manual
```{r}
clinscore = "BP_systolic_manual"
clinscore_name = "Systolic BP (manual)"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


### diastolic - manual
```{r}
clinscore = "BP_diastolic_manual"
clinscore_name = "Diastolic BP (manual)"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


### systolic - automated
```{r}
clinscore = "BP_systolic_automated"
clinscore_name = "Systolic BP (automated)"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


### diastolic - automated
```{r}
clinscore = "BP_diastolic_automated"
clinscore_name = "Diastolic BP (automated)"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## Sleep duration
```{r}
clinscore = "sleep_duration"
clinscore_name = "Sleep duration"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## Alcohol frequency
```{r}
clinscore = "Alcohol_frequency"
clinscore_name = "Alcohol frequency"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## Seen GP for depression
```{r}
clinscore = "seen_gp_depression"
clinscore_name = "Depression - GP consultation"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## Seen psychiatry for depression
```{r}
clinscore = "seen_psych_depression"
clinscore_name = "Depression - psychiatry consultation"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## Overall health rating
```{r}
clinscore = "overall_health_rating"
clinscore_name = "Overall health rating"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## Hearing difficulties
```{r}
clinscore = "hearing_difficulties"
clinscore_name = "Hearing difficulty"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```


## Neuroticism score
```{r}
clinscore = "neuroticism_score"
clinscore_name = "Neuroticism score"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(normal(50, 50), class = "Intercept"), 
            prior(normal(50, 50), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(cog_brms_draws$b_Intercept) * 1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

